---
title: "Seamless task switching ERPs: project comparison"
author: "Jussi Korpela"
date: "June 26, 2018"

output: 
  html_document:
    toc: true
    theme: united
    code_folding: hide
    
params:
  r_base_dir: /home/jkor/work_local/projects/seamless/github/R
  project_root_path: /ukko/projects/SeamlessCare_2015-16/
  
---

```{r setup}
# Set up RMarkdown working directory & some options

# Working directory for easy sourcing of helper functions
R_BASE_DIR <- params$r_base_dir
knitr::opts_knit$set(root.dir = normalizePath(R_BASE_DIR))

# Set general options
knitr::opts_knit$set(echo = TRUE,
                     fig.align = "center")
```

# Objective

Plot ERPs and related infromation. Plot extracted amplitudes and latencies (features) as well. A preliminary analysis to see if the are any effects in the data.


# TODO

* Double check how peak averaging windows etc. have been defined? Same windows for both projects or not?

* LME4 analyses: MBI cutoff (1.5 / 2.5)


# Setup
Configurations for this analysis. To see them, tap open the code block.

```{r initialize, message=FALSE}
## Setup
source("init_jkor.R")
source("tools_dataload.R")

require(cowplot)
#require(plotly)
require(DT)

MBI_CUTOFF <- 1.49 #cutoff value for MBI based grouping

FIG_WIDTH <- 15
OUT_WIDTH_FC <- 15*100

PROJECT_ROOT_PATH <- params$project_root_path
PATHS <- create_ctap_paths(PROJECT_ROOT_PATH, 'dummy')
SEAM_INPUT_PATH <- file.path(PATHS$project_root, 'analysis', 'ctap_tswitch_results', 'rds')
UUPU_INPUT_PATH <- file.path(PATHS$project_root, 'analysis', 'ctap_tswitch_uupu_results', 'rds')
SBJD_INPUT_PATH = file.path(PATHS$project_root, 'data', 'taustatiedot')

```

Setup summary:

* dataset 1: `r SEAM_INPUT_PATH`

* dataset 2: `r UUPU_INPUT_PATH`

* MBI cutoff for grouping: `r MBI_CUTOFF`


# Data


## Subject related metadata

```{r}
sbj_factor_file = file.path(SBJD_INPUT_PATH, 'uupu_seam_metadata.rds')
sbjf = as_tibble(readRDS(sbj_factor_file))

# aggregate_subject_features() comes from tools_dataload.R
mbiage <- aggregate_subject_features(sbjf, mbi_th = MBI_CUTOFF)

sbjf = sbjf %>%
  rename(sbjnr = subjnr) %>%
  mutate(sbjnr = as.integer(sbjnr))
```


```{r}
mbiage %>% group_by(mbigroup) %>% count()
```

Add mbigroup to subject data.
```{r}
sbjf <- left_join(sbjf,
                  mbiage %>% select(sbjnr, mbigroup),
                  by = 'sbjnr')
```


There is one missing data point in key subject related variables:
```{r}
sum(is.na(sbjf %>% select(age, MBIScore, BDI2, BAI)))
```

Impute missing values using group averages:
```{r}
# Replace with group mean
my_imputation <- function(ds){
  cvar <- 'BDI2'
  na.match <- is.na(ds[[cvar]])
  ds[na.match] <- mean(ds[[cvar]], na.rm = T)
  ds
}

sbjf <- sbjf %>%
  group_by(project, mbigroup) %>%
  do(my_imputation(.)) %>%
  ungroup()
```



## ERP feature data
The raw data consists of subject average ERPs stored as HDF5 files by the CTAP analysis pipe. These are collected into an R -dataset by a separate script `tswitch_prepare_erpfeat.R`. Also ERP features such as amplitude and latency are extracted by the same script and stored to disk for fast access. 

Let's first load all this data for both projects and combine.

```{r}
erpf1 = readRDS(file.path(SEAM_INPUT_PATH, 'TS03_erpfeat.rds'))
erpf1$project = 'SEAM'
erpf2 = readRDS(file.path(UUPU_INPUT_PATH, 'TS03_erpfeat.rds'))
erpf2$project = 'UUPU'

# Combine + add subject number (to tell apart cardiovasculars)
erpf = bind_rows(erpf1, erpf2) %>%
  mutate(sbjnr = as.integer(str_replace_all(subject, '[A-Z]', '')),
         project = ordered(project, levels = c('UUPU','SEAM')) )


# Combine ERP feature and subject data

## MBI group and mean age (independent of project)
erpf <- left_join(erpf,
                  mbiage %>% select(sbjnr, mbigroup, age),
                  by = 'sbjnr')

## MBIScore, BDI2, BAI (vary between projects)
erpf <- left_join(erpf,
                  sbjf %>% select(subject, MBIScore, BDI2, BAI),
                  by = 'subject')

```


## Filter 

```{r}
#MBIG_LEVELS = c('low', 'high','low -> high','high -> low')
MBIG_LEVELS = c('low', 'high')
erpf <- erpf %>%
  filter(mbigroup %in% MBIG_LEVELS, sbjnr < 100) %>%
  mutate(mbigroup = ordered(mbigroup, levels = MBIG_LEVELS ))
```

There are `r sum(is.na(erpf))` missing values in `erpf`.


Rename columns to match MMN analysis such that MMN LME models can be applied direclty:

```{r}
erpf <- erpf %>%
  rename(subjectid = subject) %>%
  mutate(subject = sbjnr)
```


## Aggregate feature values (for plotting)

```{r}
erpagg <- erpf %>%
  group_by(project,mbigroup,erpid,comp,channel,variable) %>%
  summarise(n = n(), mean = mean(value), sd = sd(value), se = sd / sqrt(n))
```

# Visualize

### Amplitude - P3a
```{r, out.width = OUT_WIDTH_FC}

my_interaction_plot <- function(pd){
  p <- ggplot(pd, aes(x = project, y = mean, color = mbigroup)) +
  geom_errorbar(aes(ymin = mean - 1.96*se, ymax = mean + 1.96*se),
                width = 0.2) +
  geom_point() +
  geom_line(aes(group = mbigroup)) +
  scale_color_discrete() +
  theme_bw() +
  facet_grid(erpid ~ channel)
  p
}

pd <- erpagg %>%
  filter(comp == 'P3a', variable == 'amplitude', mbigroup %in% c('low','high'))
p <- my_interaction_plot(pd)
p
```


### Amplitude - P3b
```{r, out.width = OUT_WIDTH_FC}
pd <- erpagg %>%
  filter(comp == 'P3b', variable == 'amplitude', mbigroup %in% c('low','high'))
p <- my_interaction_plot(pd)
p
```


### Latency - P3a
```{r, out.width = OUT_WIDTH_FC}
pd <- erpagg %>%
  filter(comp == 'P3a', variable == 'latency', mbigroup %in% c('low','high'))
p <- my_interaction_plot(pd)
p
```


### Latency - P3b
```{r, out.width = OUT_WIDTH_FC}
pd <- erpagg %>%
  filter(comp == 'P3b', variable == 'latency', mbigroup %in% c('low','high'))
p <- my_interaction_plot(pd)
p
```






# Linear Mixed Models (LMMs)


```{r}

erpfn <- erpf %>%
  group_by(erpid, comp, variable) %>%
  nest()

#erpfn[1,'data'][[1]][[1]]


require(lme4)
require(lmerTest)
source("mmn/lmm_models.R")
source("tswitch/tswitch_lmm_models.R")

# cmodel <- lmm_ia_chbdibai()
# cdata <- as.data.frame(erpfn[1,'data'][[1]])
# res <- eval_model(cmodel, cdata, p_method = "satterthwaite")

# cmodf <- function(ds){
#   eval_model(lmm_ia_chbdibai(), ds, p_method = "satterthwaite")
# }
# 
# modres <- map(erpfn$data, cmodf)
# erpfn$modres <- modres

# Wrap model specifications into a function that accepts the data as input and
# returns a list of results
make_model_func <- function(model_list){
  function(ds){
    eval_model(model_list, ds, p_method = "satterthwaite")
  }
}

cmodf1 <- make_model_func(lmm_ia_chbdibai())
modres1 <- map(erpfn$data, cmodf1)

cmodf2 <- make_model_func(lmm_ia_ch())
modres2 <- map(erpfn$data, cmodf2)

cmodf3 <- make_model_func(lmm_me_chbdibai())
modres3 <- map(erpfn$data, cmodf3)

cmodf4 <- make_model_func(lmm_me_ch())
modres4 <- map(erpfn$data, cmodf4)

```

## Plot goodness of fit (GOF) values
```{r}
make_gof_pd <- function(erpfn, modres){
  gof_pd <- bind_cols(erpfn %>% select(-data),
                    tibble(a = map(modres, function(el){el$gof})) %>% unnest())
  gof_pd <- gof_pd %>%
    gather('gof', 'value', one_of(c(names(modres[[1]]$gof))) )
  gof_pd$model <- modres[[1]]$model$description
  gof_pd
}

gof_pd <- bind_rows(
  map(list(modres1, modres2, modres3, modres4),
          function(lel){make_gof_pd(erpfn, lel)})
  )
head(gof_pd)

```

### P3a - amplitude
```{r, out.width = OUT_WIDTH_FC}
my_gof_plot <- function(pd){
  p <- ggplot(pd, aes(x = model, y = value, fill = model)) +
  geom_col() +
  scale_color_discrete() +
  theme_bw() +
  facet_grid(gof ~ erpid, scales = 'free_y') +
  theme(axis.text.x = element_text(angle = -45, hjust = 0),
        legend.position = "none")
  p
}

pd <- gof_pd %>% filter(comp == 'P3a', variable == 'amplitude')
p <- my_gof_plot(pd)
p
```

### P3b - amplitude
```{r, out.width = OUT_WIDTH_FC}
pd <- gof_pd %>% filter(comp == 'P3b', variable == 'amplitude')
p <- my_gof_plot(pd)
p
```

### P3a - latency
```{r, out.width = OUT_WIDTH_FC}
pd <- gof_pd %>% filter(comp == 'P3a', variable == 'latency')
p <- my_gof_plot(pd)
p
```

### P3b - latency

```{r, out.width = OUT_WIDTH_FC}
pd <- gof_pd %>% filter(comp == 'P3b', variable == 'latency')
p <- my_gof_plot(pd)
p
```



### Plot significant components
```{r}
ALPHA <- 0.05

make_fsig_pd <- function(erpfn, modres){
  fsig_pd <- erpfn %>%
    select(-data) %>%
    mutate(fsig = map(modres, function(el){el$fsig}),
           model = modres[[1]]$model$description) %>%
    unnest(fsig)
  fsig_pd
}

tmp <- make_fsig_pd(erpfn, modres2)

fsig_pd <- bind_rows(
  map(list(modres1, modres2, modres3, modres4),
          function(lel){make_fsig_pd(erpfn, lel)})
  ) %>%
  mutate(is_signif = p < ALPHA)

head(fsig_pd)

```


### Amplitude - P3a
```{r, out.width = OUT_WIDTH_FC}

my_interaction_plot2 <- function(pd){
  p <- ggplot(pd, aes(x = project, y = mean, color = mbigroup)) +
  geom_errorbar(aes(ymin = mean - 1.96*se, ymax = mean + 1.96*se),
                width = 0.2) +
  geom_point() +
  geom_line(aes(group = mbigroup)) +
  scale_color_discrete() +
  theme_bw() +
  facet_grid(channel ~ erpid)
  p
}

pd <- erpagg %>%
  filter(comp == 'P3a', variable == 'amplitude', mbigroup %in% c('low','high'))
p <- my_interaction_plot2(pd)
p
```

```{r, out.width = OUT_WIDTH_FC}
my_fsig_plot <- function(pd){
  p <- ggplot(pd, aes(x = model, y = p, color = is_signif)) +
    geom_point() +
    scale_color_discrete() +
    theme_bw() +
    scale_y_continuous(limits = c(0,1)) +
    facet_grid(fname ~ erpid, scales = 'free_y') +
    theme(axis.text.x = element_text(angle = -45, hjust = 0),
          legend.position = "none")
  p
}

pd <- fsig_pd %>% filter(comp == 'P3a', variable == 'amplitude')
p <- my_fsig_plot(pd)
p
```

```{r, out.width = OUT_WIDTH_FC}
pd <- erpagg %>%
  filter(comp == 'P3a', variable == 'amplitude', mbigroup %in% c('low','high'))
p <- my_interaction_plot(pd)
p
```


### Amplitude - P3b
```{r, out.width = OUT_WIDTH_FC}
pd <- erpagg %>%
  filter(comp == 'P3b', variable == 'amplitude', mbigroup %in% c('low','high'))
p <- my_interaction_plot2(pd)
p
```

```{r, out.width = OUT_WIDTH_FC}
pd <- fsig_pd %>% filter(comp == 'P3b', variable == 'amplitude')
p <- my_fsig_plot(pd)
p
```

```{r, out.width = OUT_WIDTH_FC}
pd <- erpagg %>%
  filter(comp == 'P3b', variable == 'amplitude', mbigroup %in% c('low','high'))
p <- my_interaction_plot(pd)
p
```





### Latency - P3a
```{r, out.width = OUT_WIDTH_FC}
pd <- erpagg %>%
  filter(comp == 'P3a', variable == 'latency', mbigroup %in% c('low','high'))
p <- my_interaction_plot2(pd)
p
```

```{r, out.width = OUT_WIDTH_FC}
pd <- fsig_pd %>% filter(comp == 'P3a', variable == 'latency')
p <- my_fsig_plot(pd)
p
```

```{r, out.width = OUT_WIDTH_FC}
pd <- erpagg %>%
  filter(comp == 'P3a', variable == 'latency', mbigroup %in% c('low','high'))
p <- my_interaction_plot(pd)
p
```





### Latency - P3b
```{r, out.width = OUT_WIDTH_FC}
pd <- erpagg %>%
  filter(comp == 'P3b', variable == 'latency', mbigroup %in% c('low','high'))
p <- my_interaction_plot2(pd)
p
```

```{r, out.width = OUT_WIDTH_FC}
pd <- fsig_pd %>% filter(comp == 'P3b', variable == 'latency')
p <- my_fsig_plot(pd)
p
```

```{r, out.width = OUT_WIDTH_FC}
pd <- erpagg %>%
  filter(comp == 'P3b', variable == 'latency', mbigroup %in% c('low','high'))
p <- my_interaction_plot(pd)
p
```











