---
title: "Seamless task switching ERPs: descriptive analysis"
author: "Jussi Korpela"
date: "June 26, 2018"

output: 
  html_document:
    toc: true
    theme: united
    code_folding: hide
    
params:
  r_base_dir: /home/jkor/work_local/projects/seamless/github/R
  project_root_path: /ukko/projects/SeamlessCare_2015-16/
  ctap_root_id: ctap_tswitch_uupu
  
---

```{r setup}
# Set up RMarkdown working directory & some options

# Working directory for easy sourcing of helper functions
R_BASE_DIR <- params$r_base_dir
knitr::opts_knit$set(root.dir = normalizePath(R_BASE_DIR))

# Set general options
knitr::opts_knit$set(echo = TRUE,
                     fig.align = "center")

```

# Objective

Plot ERPs and related infromation. Plot extracted amplitudes and latencies (features) as well. A preliminary analysis to see if the are any effects in the data.


# TODO


# Setup
Configurations for this analysis. To see them, tap open the code block.

```{r initialize, message=FALSE}
## Setup
source("init_jkor.R")
source("tools_dataload.R")

require(cowplot)
#require(plotly)
require(DT)


PROJECT_ROOT_PATH <- params$project_root_path
CTAP_ROOT_ID <- params$ctap_root_id
PATHS <- create_ctap_paths(PROJECT_ROOT_PATH, CTAP_ROOT_ID)
#PATHS_LOCAL <- create_ctap_paths('/home/jkor/Dropbox/Shared/seamless_jkor/ukonkuva', CTAP_ROOT_ID)

MBI_CUTOFF <- 1.49

FIG_WIDTH <- 15
OUT_WIDTH_FC <- 15*100

```

Setup summary:

* project: `r CTAP_ROOT_ID`

* MBI cutoff for grouping: `r MBI_CUTOFF`


# Data

The raw data consists of subject average ERPs stored as HDF5 files by the CTAP analysis pipe. These are collected into an R -dataset by a separate script `tswitch_prepare_erpfeat.R`. Also ERP features such as amplitude and latency are extracted by the same script and stored to disk for fast access. 

Let's first load all this data. We also split `casename` into several columns for practicality.

```{r}

savedir <- PATHS$data_root

# subject average ERPs
savefile <- file.path(savedir, 'TS03_erp_sbjave.rds')
erp_curves <- readRDS(savefile)
# split casename into a set of classifier variables
erp_curves <- erp_curves %>%
  separate(casename , c('subject','part','session','measurement'),
           remove = F) %>%
  mutate(sbjnr = as.integer(
                 str_replace_all(subject, '[A-Z]*', '')
                 ))

# ERP features
savefile <- file.path(savedir, 'TS03_erpfeat.rds')
erpfeat <- readRDS(savefile)

```


Filter out unwanted subjects:

* `sbjnr` is less than 100 for seamless: burnout study
* todo for uupuneet

```{r, filter}

if(CTAP_ROOT_ID == 'ctap_tswitch'){
  
  erp_curves <- erp_curves %>%
    filter(sbjnr < 100)
  
  erpfeat <- erpfeat %>%
    filter(sbjnr < 100)
  
}

```



## Subject average ERPs
These are averages of all available trials for a given subject and ERP. The raw ERP data looks like this:
```{r}
DT::datatable(head(erp_curves))
```

The columns are:

* `channel`: electrode position according to 10-20 system

* `time`: data point latency in `ms`

* `subject`: test subject id

* `part`: the second position from casename

* `session`: the third position from casename

* `measurement`: the fourth position from casename

* `amplitude`: ERP amplitude in microvolts

* `trial_count`: ERP trial count (included for convenience)

* `erpid`: ERP identification id

* `sbjnr`: test subject number

Let's look at the how many different levels there are for `erpid` and `channel`. The following are data row counts for different values of the classifying variable:

```{r}
erp_curves %>% group_by(erpid) %>% count()
``` 

```{r}
erp_curves %>% group_by(channel) %>% count()
``` 

There are in total `r length(unique(erp_curves$subject))` different measurements included as identified by `subject`.



## ERP features
The ERP feature data has been extracted from the subject average ERPs. It looks like this:
```{r}
DT::datatable(head(erpfeat))
```

The columns are:

* `erpid`: ERP identification id

* `comp`: ERP component within one ERP wave, the "peak" selected

* `channel`: electrode position according to 10-20 system

* `subject`: test subject id

* `part`: the second position from casename

* `session`: the third position from casename

* `measurement`: the fourth position from casename

* `variable`: ERP feature

* `value`: feature value

* `sbjnr`: test subject number



Let's look at how many different `comp`-`erpid` combinations we have and how many data rows there are per combination:

```{r}
erpfeat %>% 
  filter(variable == 'amplitude') %>%
  group_by(channel, comp, erpid) %>% 
  count()
``` 

So there is `P3a` and `P3b` for each ERP identified by `erpid`.

The definitions of each ERP are:

| Task | ERP id | Included stimuli | Description |
| -----|---------|-----------------|--------------|
| TS01 | CR | 110, 120 | Number task |
| TS02 | CR | 210, 220 | Letter task |
| TS03 | CR | 111-119, 121-129, 211-219, 221-229 | Switch task: trials at all positions |
| TS03 | CRp1 | 111, 121, 211, 221 | Switch task: switch trials position 1|
| TS03 | CRp2 | 112, 122, 212, 222 | Switch task: switch trials position 2 |
| TS03 | CRp2to9 | 112-119, 122-129, 212-219, 222-229 | Switch task: repetition trials positions 2-9|
| TS03 | CRp3to9 | 113-119, 123-129, 213-219, 223-229 | Switch task: repetition trials positions 3-9 |



# MBI scores

Here we load MBI scores for each subject and add two classification variables: a two-level and a three-level MBI classifier.
```{r}
sbjd <- load_subject_features(PROJECT_ROOT_PATH)

# Return three-level MBI class for a scalar MBI value
get_mbi_3class_single <- function(x){
  
  if (x <= 1.49){
    'low'
  } else if ((1.49 < x) & (x <= 3.49)){
    'mid'
  } else { 
    'high' 
    }
}

# Return three-level MBI class for a vector of MBI values
# To be used with mutate().
get_mbi_3class <- function(mbi_vec){
  sapply(mbi_vec, get_mbi_3class_single)
}

# Add class factors
sbjd <- sbjd %>%
  mutate(mbi_class = ordered(MBIScore > MBI_CUTOFF,
                             levels = c(T,F),
                             labels = c('high','low')),
         mbi_class3 = ordered(get_mbi_3class(MBIScore),
                              levels = c('high','mid','low'))
         )

# Save for reference
sbj_info_dir <- file.path(PROJECT_ROOT_PATH,'data','taustatiedot')
sbj_info_file <- file.path(sbj_info_dir, 'uupu_seam_metadata.xlsx')
require(xlsx)
xlsx::write.xlsx2(as.data.frame(sbjd),
                  sbj_info_file,
                  sheetName = 'data',
                  col.names = T, row.names = F, append = F)

```

The table `sbjd` contains information on both UUPU and SEAMLESS subjects. The data as a searchable table looks like this:
```{r}
DT::datatable(sbjd)
```



# Trial counts

Let's see what kind of trial counts exist in the data:
```{r trial_counts, out.width = OUT_WIDTH_FC}

ds <- erp_curves %>%
  filter(channel == 'Pz', time == 0)
p <- ggplot(ds) +
  geom_dotplot(aes(x = trial_count), binwidth = 5) +
  facet_grid(erpid ~ .) +
  scale_x_continuous(breaks = seq(0, 600, 50)) +
  theme_linedraw() +
  labs(x = 'Subject trial count for ERP', y = 'Count of subjects')
p
```

Most of these seem to be above 50 which should suffice for a large ERP such as the P3.

A zoomed-in version of the figure above:

```{r trial-counts-zoom, out.width = OUT_WIDTH_FC}
p <- ggplot(ds) +
  geom_dotplot(aes(x = trial_count), binwidth = 5) +
  facet_wrap(erpid ~ ., scales = 'free_x', ncol = 2) +
  theme_linedraw() +
  labs(x = 'Subject trial count for ERP', y = 'Count of subjects')
p

```



Let's also see if the high MBI score subjects have also low trial counts? This is suspected as their ERPs will later prove to be small...

```{r trial_counts_byMBI, out.width = OUT_WIDTH_FC}
ds <- erp_curves %>%
  filter(channel == 'Pz', time == 0)

# get MBI classes
ds <- left_join(ds,
                select(sbjd, c('subject', 'mbi_class', 'mbi_class3', 'MBIScore')),
                by = 'subject')

p <- ggplot(ds) +
  geom_point(aes(x = trial_count, y = MBIScore, colour = mbi_class3)) +
  facet_wrap(erpid ~ ., scales = 'free_x') +
  theme_light() +
  scale_colour_brewer(palette = 'Set1') +
  labs(x = 'Subject trial count for ERP', y = 'MBI score')
p


```

The answer is no: high MBI subjects have decent trial counts.



# Grand Average (GA) ERP

## GA ERP for all subjects
The grand average is the average ERP averaged over all test subjects. Here are grand average ERPs for each erpid separately:
```{r, fig.height = 20, fig.width = FIG_WIDTH}
my_gaerp_plot <- function(pd){

  cur_erpid <- unique(pd$erpid)

  pd <- pd %>%
    mutate(value = amplitude,
           ds = subject)

  gapd <- pd %>%
    group_by(channel, time) %>%
    summarise(n=n(), mean=mean(amplitude)) %>%
    ungroup()

  titlestr = sprintf('%s GA ERP, N_sbj = %d', cur_erpid, unique(gapd$n))
  p <- ggplot.gaerp(gapd,
                     ylimits = c(15, -7),
                     titlestr = titlestr) #plot
  p

  # gp <- ggplotly(p)
  # savedir_loc <- file.path(RESULTS_LOCAL_BASEPATH, 'figs', 'GA')
  # dir.create(savedir_loc, showWarnings = F, recursive = T)
  # savefile <- file.path(savedir_loc, sprintf('GAERP_%s.html', cur_erpid))
  # htmlwidgets::saveWidget(gp, savefile)
  # note: using local save directory as saving to Ukko fails
  # for unknown reason.

  #tibble(success = T, plot = list(p))
}

erpid_arr <- unique(erp_curves$erpid)
p_list <- vector(length(erpid_arr), mode = 'list')
i = 1
for (id in erpid_arr){
  p <- my_gaerp_plot(filter(erp_curves, erpid == id))
  p_list[i] <- list(p)
  i <- i +1
}
cowplot::plot_grid(plotlist = p_list, ncol = 2)
```

The two components are clearly visible. Because the ERPs are visual, we use channel `Pz` for finding the GA peak position and for extracting the ERP features.


## GA ERP by MBI group

```{r}
# Merge MBI information into ERP curve data

mbiage <- aggregate_subject_features(sbjd, mbi_th = MBI_CUTOFF)

erp_curves_mbi <- left_join(erp_curves,
                            mbiage %>% select(sbjnr, mbigroup),
                            by = 'sbjnr')

erp_curves_mbi <- left_join(erp_curves_mbi,
                            sbjd %>% select(subject, mbi_class, mbi_class3),
                            by = 'subject')
```


### MBI grouping - project specific

```{r, fig.height = 15, fig.width = FIG_WIDTH}

my_gaerp_plot_bymbi <- function(gapd, class_variable){

  ext.val <- max(abs(gapd$mean_amp))
  ylimits <- c(ceiling(ext.val), -ceiling(ext.val))
  
  p <- ggplot(data = gapd) +
    geom_line(aes_string(x = 'time', y = 'mean_amp',
                  group = class_variable, color = class_variable)) +
    geom_hline(yintercept = 0, size = 0.3) +
    geom_vline(xintercept = 0, size = 0.3) +
    facet_grid(erpid ~ channel) +
    scale_colour_brewer(type = 'qual', palette = 2) +
    scale_y_reverse(limits = c(15, -7)) + #should apply to both projects!
    labs(x = 'Time (ms)', y = 'Amplitude (muV)')
  p
}


gapd1 <- erp_curves_mbi %>%
  group_by(part, session, mbi_class, erpid, channel, time) %>%
  summarise(n = n(), mean_amp = mean(amplitude)) %>%
  ungroup() %>%
  mutate(mbi_class = fct_rev(mbi_class))

p <- my_gaerp_plot_bymbi(gapd1, 'mbi_class') +
  labs(title = 'GA ERP by project MBI grouping, 2 levels')
p
```


```{r, fig.height = 15, fig.width = FIG_WIDTH}
gapd2 <- erp_curves_mbi %>%
  group_by(part, session, mbi_class3, erpid, channel, time) %>%
  summarise(n = n(), mean_amp = mean(amplitude)) %>%
  ungroup() %>%
  mutate(mbi_class3 = fct_rev(mbi_class3))

p <- my_gaerp_plot_bymbi(gapd2, 'mbi_class3') +
  labs(title = 'GA ERP by project MBI grouping, 3 levels')
p
```


### MBI grouping - independent of project

```{r, fig.height = 15, fig.width = FIG_WIDTH}
gapd3 <- erp_curves_mbi %>%
  group_by(part, session, mbigroup, erpid, channel, time) %>%
  summarise(n = n(), mean_amp = mean(amplitude)) %>%
  ungroup()

p <- my_gaerp_plot_bymbi(gapd3 %>% filter(mbigroup %in% c('low','high')),
                         'mbigroup') +
  labs(title = 'GA ERP project-independent MBI grouping')
p
```



# ERP features

First we add the MBI data to ERP feature data.

```{r}
# Merge into ERP feature data
pd <- left_join(erpfeat,
                select(sbjd, c('subject', 'mbi_class', 'mbi_class3', 'MBIScore')),
                by = 'subject') %>%
  mutate(erpid_mbi = paste0(erpid,'_',mbi_class))

```


## Boxplot using two MBI classes
Here a two-level MBI classification is used. Cutoff at MBIScore == `r sprintf('%1.1f', MBI_CUTOFF)`.

```{r erpfeat_boxplot_2class, fig.height = 20, out.width = OUT_WIDTH_FC}
## Boxplot

erpid_boxplot <- function(pds, mbi_class_var){
  p <- ggplot(filter(pds, variable == 'amplitude')) +
    geom_boxplot(aes_string(x = mbi_class_var, y = 'value',
                     color = mbi_class_var)) +
    geom_jitter(aes_string(x = mbi_class_var, y = 'value'),
                alpha = 0.25, size = 0.5, width = 0.25) +
    facet_grid(comp ~ channel) +
    theme_light() +
    theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
    scale_colour_brewer(palette = "Set1") +
    labs(x = 'Target stimulus', y = 'Amplitude (uV)',
         title = sprintf('%s ERP amplitudes within MBI groups', unique(pds$erpid)))
  p
}

erpid_arr <- unique(erpfeat$erpid)
p_list <- vector(length(erpid_arr), mode = 'list')
i = 1
for (id in erpid_arr){
  p <- erpid_boxplot(pd %>% 
                       filter(erpid == id),
                     'mbi_class')
  p_list[i] <- list(p)
  i <- i +1
}
cowplot::plot_grid(plotlist = p_list, ncol = 1)

#tmp <- pd %>% filter(erpid == id)
#erpid_boxplot(pd %>% filter(erpid == id), 'mbi_class')

```


## Boxplot using three MBI classes
Here a three-level MBI classification is used. Cutoffs at 1.49 and 3.49 as in Ahola et. al. (2005).

Sokka et. al. (2017) use also three groups but the MBI cutoffs are hard to find from that paper. However, Ahola et. al. (2005) define them as follows:

"Burnout and the dimensional scores were categorized as follows: No burnout (scores 0–1.49), mild burnout (scores 1.50–3.49), and severe burnout (scores 3.50–6)."

Reference: Ahola, K., Honkonen, T., Isometsä, E., Kalimo, R., Nykyri, E., Aromaa, A., & Lönnqvist, J. (2005). The relationship between job-related burnout and depressive disorders—results from the Finnish Health 2000 Study. Journal of affective disorders, 88(1), 55-62.


```{r erpfeat_boxplot_3class, fig.height = 20, out.width = OUT_WIDTH_FC}
erpid_arr <- unique(erpfeat$erpid)
p_list <- vector(length(erpid_arr), mode = 'list')
i = 1
for (id in erpid_arr){
  p <- erpid_boxplot(pd %>% 
                       filter(erpid == id),
                     'mbi_class3')
  p_list[i] <- list(p)
  i <- i +1
}
cowplot::plot_grid(plotlist = p_list, ncol = 1)
```

# Parallel coordinates plots
Here is a parallel coordinates plot showing that (todo: only true fro seam):

* high MBI subjects have very similar low amplitude ERPs accross erpid:s

* some low MBI subjects show similar behavior

```{r, out.width = OUT_WIDTH_FC}
## Parallel coordinates plot
pdtmp <- pd %>%
  mutate(comp_mbi = paste0(comp,'_',mbi_class),
         comp_mbi3 = paste0(comp,'_',mbi_class3)) %>%
  filter(erpid %in% c('CR','CRp1','CRp2','CRp3to9'))
p <- ggplot(filter(pdtmp, variable == 'amplitude')) +
  geom_line(aes(x = erpid, y = value, color = mbi_class3, group = subject)) +
  geom_point(aes(x = erpid, y = value, color = mbi_class3, group = subject),
             size = 0.5) +
  facet_grid(channel ~ comp_mbi3) +
  scale_color_brewer(palette = 'Set1') +
  theme_light() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  labs(x = 'Target stimulus', y = 'Amplitude (uV)',
       title = 'ERP amplitudes within MBI groups')
p
```




